---
title: "My R Markdown Primer"
author: "Animesh Singh"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

R Markdown is a tool that combines regular text with R code to create dynamic documents. You write your analysis and narrative together, then "knit" it into a polished HTML, PDF, or Word document. This makes research reproducible since code and output stay together in one file.

## Math Expressions

R Markdown supports LaTeX math. Here's an inline example: $\alpha + \beta = \gamma$

And here's a displayed equation for the sample mean:

$$
\bar{x} = \frac{1}{n}\sum_{i=1}^{n}x_i
$$

## Loading Data

```{r load-data}
library(ggplot2)

customers <- read.csv("/Users/animeshsingh/Documents/r-programming-assignments-main/r-programming-assignments/data/customers-100.csv")
head(customers, 5)
```

## Simple Analysis

```{r summary}
# Basic summary statistics
cat("Total customers:", nrow(customers), "\n")
cat("Number of countries:", length(unique(customers$Country)), "\n")

# Top 5 countries
country_counts <- table(customers$Country)
top5 <- sort(country_counts, decreasing = TRUE)[1:5]
print(top5)
```

## Visualization

```{r plot, fig.width=8, fig.height=5}
# Convert subscription date and extract year
customers$Subscription.Date <- as.Date(customers$Subscription.Date)
customers$Year <- format(customers$Subscription.Date, "%Y")

# Plot subscriptions by year
year_counts <- table(customers$Year)
barplot(year_counts, 
        main = "Customer Subscriptions by Year",
        xlab = "Year", 
        ylab = "Number of Customers",
        col = "steelblue")
```

## Reflection

**R Markdown vs. Plain Reports:**

R Markdown is better than traditional reports because:

- Code and results stay together: no copy/paste errors
- Easy to update when data changes: just re-knit
- Fully reproducible: anyone can verify your work

The main challenge is learning the syntax, but it's worth it for the time saved and improved reproducibility.
